A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2014-02-22 jacobian_calc_CMAME.tex
removed treatise on byte copiable data I had crammed into the conclusion, as excess

The CS method was shown to be highly accurate in calculating tangent-stiffness
matrices by the standards of this study, and it is slower than other methods
for a naive implementation. In the context of future use of CS in
\emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types,
while AD requires a complicated serialization and deserialization in order to
function as byte-copyable. This is fine for copying in between MPI processes,
but deserialization on an accelerator or GPU is very difficult, as this would
require compiling a version of the AD library being used that were compatible
with the accelerator or GPU being used. \note[Michael]{re:
I.3}\add[Michael]{For the reader's convenience, a collection of contiguously
stored data elements are said to be 'byte-copiable' when the interval
in bytes of memory spanned by a data element in the collection is constant
throughout the collection. Suppose 'contiguously stored' means, stored in
memory with a minimum of unrelated information stored in between data
elements in the collection. Complicated objects, such as AD datatype
elements, do not have the same size in memory as other objects even of that
same class, for the reason that they themselves contain an uncertain
amount of state data according to their design and work done in the
program. This means that a collection of such objects, even if they are
contiguously stored in memory, will not be alligned according to a constant
interval, such that elements within a collection cannot be located by
coordinates consisting of the location of the beginning of the first
element in the collection and an interval multiplied by a natural number.
This method of locating is integral to the serialized communication
implemented by MPI the libraries as well as CUDA, OpenCL, and others,
which allow only the transmission of collections of byte-copiable datatypes
(that is 'primitive datatypes'). A common method of bypassing this
limitation for communication purposes is to write the complicated data to a
'character buffer', and transmit the byte-copiable character buffer,
however this is useless unless the recipient device of the communication
can recover the original data and process it, which is impossible if the
recipient device only can work with a subset of the programming language
that the sending device must use, such as in a communication between a CPU
running C++ and a GPU running C (such as in CUDA programs), where the gap in
sophistication contains the features of the "complicated" datatype.} What this means is
that if \emph{Peridigm} or any simulation software were to be reformulated to
take advantage of accelerators or GPUs, the flexibility of complex-step owed to
its simple implementation may make it more viable for that application given
its high accuracy. Additional performance gains may be achieved through the use
of the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
use of complex data types natively, and alleviate the expense of the
dynamic-casts that were utilized in this study.  As a final comment, it should
be noted that from a code development perspective the CS method may be the
easiest of the methods to implement; It only requires one functional
evaluation, there is no heuristic choice of probe distance, and there is no
dependence on an external AD library.  From this perspective, the method may
still be an attractive method for TS matrix evaluation. \add[Michael]{However, at least for 
the models producing the results shown here, each of the
methods returned quadratic convergence, indicating each was sufficiently
accurate. 'Marching orders' for the analyst would be to select the swiftest
method, while it would seem accuracy above what is adequate does not enhance
the convergence rate or iteration speed until either classical Newton's convergence
conditions or the approximate case conditions are violated.}

>>>>

The CS method was shown to be highly accurate in calculating tangent-stiffness
matrices by the standards of this study, and it is slower than other methods
for a naive implementation. In the context of future use of CS in
\emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types,
while AD requires a complicated serialization and deserialization in order to
function as byte-copyable. This is fine for copying in between MPI processes,
but deserialization on an accelerator or GPU is very difficult, as this would
require compiling a version of the AD library being used that were compatible
with the accelerator or GPU being used.  What this means is
that if \emph{Peridigm} or any simulation software were to be reformulated to
take advantage of accelerators or GPUs, the flexibility of complex-step owed to
its simple implementation may make it more viable for that application given
its high accuracy. Additional performance gains may be achieved through the use
of the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
use of complex data types natively, and alleviate the expense of the
dynamic-casts that were utilized in this study.  As a final comment, it should
be noted that from a code development perspective the CS method may be the
easiest of the methods to implement; It only requires one functional
evaluation, there is no heuristic choice of probe distance, and there is no
dependence on an external AD library.  From this perspective, the method may
still be an attractive method for TS matrix evaluation. \add[Michael]{However, at least for 
the models producing the results shown here, each of the
methods returned quadratic convergence, indicating each was sufficiently
accurate. 'Marching orders' for the analyst would be to select the swiftest
method, while it would seem accuracy above what is adequate does not enhance
the convergence rate or iteration speed until either classical Newton's convergence
conditions or the approximate case conditions are violated.}
